### 2018-04-07 book.tex

%% This file is automatically generated, DO NOT EDIT

\renewcommand{\REPORT}{THESIS}
\renewcommand{\GROUP}{None}
\renewcommand{\thepage}{}
\include{tpg}	\cleardoublepage	\newpage
\pagenumbering{roman}
\setcounter{page}{1}
\include{intro}		\newpage
\cleardoublepage
\pagenumbering{arabic}
\setcounter{page}{1}
\GEOheader{\GROUP, \REPORT, \today}
%% start of paper list
\setfigdir{Fig}\GEOpaper{chapter-locfreq}{chapter-locfreq}	\include{chapter-locfreq/paper}
\cleardoublepage
%% end of paper list
\cleardoublepage
\addcontentsline{toc}{chapter}{Bibliography}
\bibliographystyle{seg}
\bibliography{thesis}
\begin{vita}
\input{./vita}
\end{vita}