### 2018-04-07 chapter-intro/paper.tex

\title{Introduction}
\author{Sarah Yvonne Greer}
\maketitle

\subsection{Background}
Seismic data contains fundamentally non-stationary variations.