### 2018-04-07 chapter-locfreq/paper.tex

\title{Balancing local frequency content in seismic data using non-stationary smoothing}
\author{Sarah Yvonne Greer}
\maketitle

\begin{abstract}
        Seismic data can experience non-stationary frequency variations caused by attenuation. This
        problem is encountered when matching multiple data sets, such as in multicomponent image
        registration, because signals with differing frequency content are hard to correlate. In
        this paper, we propose a method to balance frequency content between data sets while taking
        into account non-stationary frequency variations. This method involves finding and applying
        a non-stationary smoothing operator to minimize the local frequency difference between data
        sets. Numerical examples demonstrate that the proposed method improves multicomponent image
        registration and matching images of differing resolution. 
\end{abstract}

\section{Introduction}
        Matching seismic data has many applications in geophysical processing methods, such as
        multi-component image registration, time-lapse image registration, matching well-ties to
        seismic data, and merging seismic data sets \cite[]{ps,fomel2003,lumley,herrera2012}.
        Typically, the workflow for matching data involves finding the optimal time shift, or amount
        of stretching and compressing, of one trace relative to the other that produces the greatest
        similarity between the two traces, as seen in dynamic time warping and local similarity
        scanning \cite[]{hale2013, timelapse, herrera}. However, when the two signals that need to
        be aligned have different spectral content, their comparison can be difficult. Most seismic
        data experience non-stationary, or spatially and temporally variant, frequency content
        caused by attenuation. This problem was discussed in application to multicomponent image
        registration by \cite{fomel2003}, who applied frequency balancing methods to improve
        registration results. \cite{ltft} proposed using local time-frequency decomposition (LTFD)
        to balance frequencies between multicomponent data during registration. However, LTFD is a
        relatively computationally expensive method.

        In this paper, we propose a method to remove non-stationary frequency differences that limit
        the effectiveness of matching data. We suggest applying the proposed method to processing
        flows that involve matching data before attempting to find the time shift to align their
        signal content. To balance frequency content, we use a non-stationary smoothing operator
        with an adjustable smoothing radius to apply to the higher frequency data set. Our approach
        takes the form of an optimization problem which is solved using an iterative method. We
        apply this method to examples of merging high-resolution and conventional seismic images and
        multicomponent image registration.

\section{Method}
\inputdir{merge}
        Two signals of differing frequencies are more difficult to correlate than signals of similar
        frequencies.  For example, Figure \ref{fig:legacy,hires-agc} shows two seismic images
        representing the same subsurface, except they have distinctly different spectral content, as
        shown in Figure \ref{fig:nspectra-orig}. In order to be directly comparable, these two
        images should have similar frequency content.
        In this paper, we look at local frequency \cite[]{attr}, which can be thought of as a
        smoothed estimate of instantaneous frequency \cite[]{white}. Local frequency is a more
        geologically accurate attribute than instantaneous frequency because it honors
        time-frequency uncertainty and does not contain physically unrealistic negative or extremely
        high frequency values \cite[]{attr}.

        In order to balance local frequency content, we propose smoothing the higher frequency data
        using a non-stationary triangle smoothing operator with an adjustable radius. Here, the
        radius at each point is the number of samples in time that that specific data point gets 
        averaged over
        %in either direction 
        in a triangle weight.

        %Our goal is to specifically remove the higher frequency signal variations

        Our goal is to find the temporally and spatially variable smoothing radius, $\mathbf{R}$, that
        minimizes the difference in local frequencies between the two data sets. This can be shown
        in the objective function
        \begin{equation}
                \label{eq:obj}
                \min_{\mathbf{R} \in [1,N]} \Big \Vert \mathbf{F}[ \mathbf{S}_{\mathbf{R}} \mathbf{d}_h ] - 
                \mathbf{F}[ \mathbf{d}_l] \Big \Vert\;,
        \end{equation}
        where $\mathbf{S}_\mathbf{R}$ is the non-stationary smoothing operator of smoothing radius
        $\mathbf{R}$, $\mathbf{d}_h$ is the higher frequency data, $\mathbf{d}_l$ is the lower
        frequency data, $\mathbf{F}$ is the local frequency operator, and $N$ is the maximum size of
        the smoothing radius.
        Although smoothing is a linear operation, the smoothed data, $\mathbf{S_R d}_h$, depends
        non-linearly on $R$. However, the objective from equation (\ref{eq:obj}) is nearly convex,
        and we choose to use an intuitive iterative approach to find an approximate smoothing
        radius.

        The main premise behind the method comes from the fact that, in general, 
        %the more samples the data are smoothed over, 
        the greater the smoothing radius, the more high frequencies are attenuated by smoothing.
        We choose an initial guess of the non-stationary smoothing radius before applying
        corrections based on two primary assumptions:
        \begin{enumerate}
        \item The smoothing radius is too \emph{small} at a specified point if, after smoothing, the
                higher frequency data has \emph{higher} local frequency than the lower frequency
                data. Thus, the smoothing radius must be \emph{increased} at that point. 
        \item The smoothing radius is too \emph{large} at a specified point if, after smoothing, the
                higher frequency data has \emph{lower} local frequency than the lower frequency
                data. Thus, the smoothing radius must be \emph{decreased} at that point.
        \end{enumerate}
        We apply these assumptions using a line-search method:
                        \begin{equation}
                               \label{eq:ls}
                               \mathbf{R}^{(i+1)} = \mathbf{R}^{(i)}+ \alpha \mathbf{r}^{(i)},
                        \end{equation}
        where $\mathbf{R}^{(i)}$ is the smoothing radius at the $i$th iteration, $\alpha$ is a
        scalar constant that can be thought of as the step length, and $\mathbf{r}^{(i)}$ is the
        residual at the $i$th iteration, which can be thought of as the search direction, and is
        defined as
          \begin{equation}
                \label{eq:residual}
                \mathbf{r} = \mathbf{F}[\mathbf{S}_{\mathbf{R}} \mathbf{d}_h] - \mathbf{F}[\mathbf{d}_l]\;.
          \end{equation}
        It can be noted that when equation (\ref{eq:residual}) is positive, the higher frequency
        data still has a higher local frequency value at that specific point than the lower
        frequency data, thus the higher frequency data is under-smoothed and the smoothing radius
        should be increased at that point. This follows the form of the first assumption. The second
        assumption is used when equation (\ref{eq:residual}) is negative. When equation
        (\ref{eq:residual}) is zero, the correct radius has been found and no further corrections
        are made. Thus, it is justifiable to set the search direction from equation (\ref{eq:ls})
        equal to the residual.

        Using the two assumptions, we can continually adjust the smoothing radius until we 
        achieve the desired result of balancing the local frequency content between the two data
        sets. In practice, this method produces an acceptable solution in approximately 5 iterations
        and exhibits sublinear convergence. After smoothing the higher frequency data with the
        estimated radius, we use the lower frequency and smoothed higher frequency data to estimate
        time shifts and align the two data sets.

\section{Examples}
        This method is applicable to workflows that require matching data with different frequency
        content. In this paper, we demonstrate two examples where smoothing data before matching
        produces better results.

        \subsection{Matching legacy and high-resolution images}
        High-resolution seismic data, such as those acquired with the P-cable acquisition system,
        can produce very detailed images of the near subsurface \cite[]{tip}. When compared to
        conventional seismic images, high-resolution images have a higher dominant frequency and a
        larger frequency bandwidth. However, they usually lack low frequency content and depth
        coverage that is present in conventional seismic images. As a result, successful
        interpretation of high-resolution images can be aided by matching with legacy data coverage
        over the same area.

        \multiplot{2}{legacy,hires-agc}{width=0.9\columnwidth}{Initial legacy (a) and
        high-resolution (b) images. These images show the same subsurface geology, but look
        remarkably different as the high-resolution image has distinctly higher frequency content
        than the legacy image.}

        \plot{rect5}{width=0.9\columnwidth}{The smoothing radius, which is a function of time and
        space, this method produces after 5 iterations. This represents the number of samples in
        time that the high-resolution image needs to be smoothed over in a triangle weight %in time
        to balance local frequency content with the legacy image.}

        Example legacy and high-resolution images of the same subsurface are shown in Figure
        \ref{fig:legacy,hires-agc}. The first step in matching the two images is to ensure that they
        both have a similar frequency bandwidth so they are directly comparable. The average frequency
        spectra for the two images are shown in Figure \ref{fig:nspectra-orig}. From this, it is
        evident that there is almost no overlap in frequency bandwidth between the two images.
        To address this problem, we apply a low-cut filter to the legacy image to remove some of the
        lower frequencies that are simply not present in the high-resolution image. Next, we
        implement the method described in the previous section to balance local frequency content
        between the two images. The difference in local frequencies (residual, by equation
        \ref{eq:residual}) between the high-resolution and legacy images before balancing frequency
        content and after 5 iterations of the algorithm in equation (\ref{eq:ls}) is shown in Figure
        \ref{fig:freqdif,freqdif-filt5}. After balancing local frequencies, the images show a
        similar spectral bandwidth (Figure \ref{fig:high-smooth-spec5}), which helps increase the
        correlation between the two images and makes matching reflections better defined.

        After the frequency content is matched, the optimal time shift is found to align signal
        content between the legacy and high-resolution images. We then apply this time shift to
        %the original legacy and high-resolution images---the frequency content is only degraded for
        the original high-resolution image---the frequency content is only degraded for the purpose
        of finding the time shift.

        An application of aligning the high-resolution and legacy images is discussed in the companion
        paper \cite[]{merge}.

        \multiplot{2}{nspectra-orig,high-smooth-spec5}{width=0.81\columnwidth}{Spectral content of
        the legacy (red) and high-resolution (blue) images before (a) and after (b) spectral
        balancing.}

\subsection{Multicomponent image registration}
        Multicomponent seismic image registration is an important step before the interpretation of
        P and S images of the subsurface. It involves warping the space of S images to align
        reflections with the analogous reflections of P images \cite[]{fomel2003,warp}.

        Figure \ref{fig:pp,ss} shows PP and SS images from a 9-component seismic survey \cite[]{attr}.
        To properly register the images, we follow the
        method proposed in \cite{warp}. It consists of three primary steps: (1) initial registration
        of PP and SS images using initial interpretation and well-log analysis; (2) balancing
        frequency and amplitude content; and (3) final registration using residual local similarity
        scanning. We incorporate our method of balancing frequency content into the second step in
        this process.

        Before initial registration, the PP image has much higher frequency content than the SS
        image. After the SS image is temporally compressed to PP time for initial registration, the
        two images have more similar frequency content. However, additional frequency balancing is
        still needed before residual registration. This poses a problem as neither image has
        distinctly higher frequencies than the other, so both images need to be smoothed in
        different areas to balance frequency content. In order to do this, we modify the proposed
        method to include two separate smoothing operators---one for each image.

        \multiplot{2}{freqdif,freqdif-filt5}{width=0.81\columnwidth}{Difference in local frequencies
        (residual) between the legacy and high-resolution images before (a) and after (b) the 5th
        iteration of frequency balancing.}

        We modify the objective in equation (\ref{eq:obj}) as
            \begin{equation}
                \label{eq:obj2}
                \min_{\mathbf{R} \in [-N, -1] \cup [1, N]} \Big \Vert \mathbf{F}[
                \mathbf{S}_{\mathbf{R}_p} \mathbf{d}_{p} ] - \mathbf{F}[
                \mathbf{S}_{\mathbf{R}_s}\mathbf{d}_{s}] \Big \Vert\;, 
            \end{equation}
        %were $S_{R_p}$ and $S_{R_s}$ are the non-stationary smoothing operators for the PP and
        %SS images, respectively, and 
        where $\mathbf{d}_p$ and $\mathbf{d}_s$ are the PP and SS images, respectively, and
        $\mathbf{S}_{\mathbf{R}_p}$ and $\mathbf{S}_{\mathbf{R}_s}$ are the non-stationary smoothing
        operators for the PP and SS images, respectively.
        We also modify the residual from equation (\ref{eq:residual}) as
            \begin{equation}
                \label{eq:residual2}
                \mathbf{r} = \mathbf{F}[\mathbf{S}_{\mathbf{R}_{p}}\mathbf{d}_p] -
                \textbf{F}[\mathbf{S}_{\mathbf{R}_s}\mathbf{d}_s]\;.
            \end{equation}
        The ideal radius is still found using the same line-search from equation (\ref{eq:ls}),
        except we allow the smoothing radius to be negative. Physically, a negative smoothing radius
        would signify that the image should be sharpened at that particular point instead of
        smoothed. In this case, instead of trying to sharpen the PP image at that particular point, 
        we
        %choose to smooth the SS image by the negative component of the smoothing radius. Thus, we
        %define $\mathbf{R}_p$ anda $\mathbf{R}_s$ at each sample as
        %\begin{equation}
        %        \mathbf{R}_p = \begin{cases} \mathbf{R} &\mbox{if } \mathbf{R} \ge 1 \\
        %1 & \mbox{otherwise} \end{cases} 
        %\end{equation}
        %and
        %\begin{equation}
        %        \mathbf{R}_s = \begin{cases} |\mathbf{R}| &\mbox{if } \mathbf{R} \le -1 \\
        %1 & \mbox{otherwise} \end{cases} 
        %\end{equation}
        %choose to smooth the SS image by the negative component of the smoothing radius. Thus, we
        %define the $i$th components of $\mathbf{R}_p$ and $\mathbf{R}_s$ as
        choose to smooth the SS image by the negative part of the smoothing radius. 
        Thus, we define the $i$th components of $\mathbf{R}_p$ and $\mathbf{R}_s$ as
        %Thus, we define $R_{p,i}$ and $R_{s,i}$ (the $i$th components of $\mathbf{R}_p$ and
        %$\mathbf{R}_s$) as
        \begin{equation}
                R_{p,i} = \begin{cases} R_{i} &\mbox{if } R_{i} \ge 1 \\
        1 & \mbox{otherwise} \end{cases} 
        \end{equation}
        and
        \begin{equation}
                R_{s,i} = \begin{cases} |R_{i}| &\mbox{if } R_{i} \le -1 \\
        1 & \mbox{otherwise} \end{cases} 
        \end{equation}

        where $R_i$ is the $i$th component of $\mathbf{R}$ and a radius of 1 represents no
        smoothing. This allows each image to be smoothed in different areas to balance the frequency
        content between the two images.

\inputdir{vecta}
        \multiplot{2}{pp,ss}{width=0.85\columnwidth}{Initial PP (a) and SS (b) images.}

        The results of using this spectral balancing method are shown in Figure
        \ref{fig:before,after}. Comparable results were achieved in \cite{ltft}, who used LTFD to
        balance the spectral content between the two images. However, the method we propose in this
        paper is more straightforward and significantly less computationally expensive.

\section{Conclusions}
        It is difficult to compare signals with differing frequency content because signals with
        differing frequencies are hard to correlate. In this paper, we proposed a method of
        balancing frequency content between data which takes the form of an optimization problem
        solved by a simple iterative algorithm. It is a relatively inexpensive and effective method
        compared to previously proposed methods of balancing frequencies between data sets. We
        applied this method to examples of matching seismic images of differing resolution and for
        multicomponent image registration.

        \multiplot{2}{before,after}{width=0.85\columnwidth}{Interleaved PP and warped SS traces before
        (a) and after (b) frequency balancing and residual registration. After registration, the
        signal content between the two initial images is more aligned; for example, the reflections
        around 0.3 and 0.6 s. This indicates a successful registration.}

\section{Acknowledgements}
        We thank the sponsors of the Texas Consortium for Computational Seismology (TCCS) for their
        financial support.

\onecolumn

%\bibliographystyle{seg}
%\bibliography{SEG,locfreq}