### 2018-04-22 ack.tex
almost done with thesis git st

I have received an incredible amount of support during my time as an undergraduate student. 
First and foremost, I'd like to thank my advisor, Dr. Sergey Fomel, for his helpful guidance over the past couple of years. 
I genuinely believe that working with him has beneficially shaped the trajectory of my life.
Through his guidance and the opportunities he has given me, he has helped me to realize how much I truly enjoy learning, showed me the importance of reproducible computational research, introduced me new ways of thinking about and approaching problems, and inspired me to continue my studies in graduate school.
For all that, I will be forever grateful.

Many professors from the geological sciences and mathematics departments with whom I have interacted have also greatly helped me during my undergraduate career.  
I am especially thankful to Dr. Todd Arbogast, Dr. Jaime Barnes, Dr. Mark Cloos, Dr. Luc Lavier, Dr. Francesco Maggi, Dr. Richard Tsai, and my committee members, Dr. Kyle Spikes and Dr. Clark Wilson, for their guidance during my time at UT Austin.

I am also grateful to many students and research scientists I've had the pleasure of interacting with: 
Sean Bader, Luke Decker, Preston Fussee-Durham, Zhicheng Geng, Eric Goldfarb, Ben Gremillion, Tom Hess, Ken Ikeda, Harpreet Kaur, Dr. Dmitrii Merzlikin, Nam Pham, Mason Phillips, Kelly Regimbal, Karl Schleicher, Yunzhi Shi, Dr. Junzhe Sun, Dr. Yanadet Sripanich, Dr. Xinming Wu, Dr. Zhiguang Xue, and Dr. Yunan Yang.

Last but not least, I would like to express my sincere gratitude to my family for all their love and support over the years.
Esme, Joel, Sam, Atticus, and Kali---you all mean the world to me and I wouldn't be the person I am today without you in my life.
\begin{flushright}
SARAH YVONNE GREER
\end{flushright}

\noindent \emph{The University of Texas at Austin}\\
\noindent \emph{April 2018} \\ \\