### 2018-04-22 vita.tex

Sarah graduated with high honors from Cinco Ranch High School in Katy, Texas in 2014 and came to The University of Texas at Austin (UT Austin) to pursue her undergraduate studies. 
%She is a candidate for two Bachelor of Science degrees from UT Austin in Geological Sciences and Mathematics. 
She is a candidate for a Bachelor of Science degree in Geophysics from the Jackson School of Geosciences and a Bachelor of Science degree in Mathematics from the College of Natural Sciences.
After graduation, she plans on pursuing a PhD in Mathematics and Computational Science from the Massachusetts Institute of Technology.

%After graduation, she plans on pursuing a PhD in Computational Science, Engineering, and Mathematics from UT Austin's Institute for Computational Engineering and Sciences.
%After graduation, she plans on pursuing a PhD in Computational and Mathematical Engineering from Stanford University.