### 2018-04-23 abs.tex
done... can't get thesis to compile though!

Data matching applications are present in many computational geophysics problems.
In this thesis, I introduce a new algorithm to aid in data matching that balances local frequency content between seismic data sets.
Then, I provide a few examples where applying this algorithm to seismic data helps improve results.
These applications include multi-component image registration, matching and merging high-resolution and legacy seismic images, and improving migration resolution by approximating the least-squares Hessian.
The applications are applied to several 2D and 3D, real and synthetic data sets.