### 2018-05-03 abs.tex
thesis almost done

\author{Sarah Greer, The University of Texas at Austin}
\title{Abstract}
\label{ch:abs}

Data matching applications appear in many computational geophysics problems.
In this thesis, I introduce a new efficient algorithm to aid in data matching that balances local frequency content between seismic data sets.
Then, I provide a few examples where applying this algorithm to seismic data helps improve results.
These applications include multi-component image registration, matching and merging high-resolution and legacy seismic images, and improving migration resolution by approximating the least-squares Hessian.
The applications are illustrated by several 2D and 3D, real and synthetic data sets.